Thesis: draft

### 2021-04-11 draft.tex
added extra result

From Figure \ref{ev_op_20_80}, it is clear that the minority opinion 0 eventually ceases to exist for all the network models, $P_0(500) \sim 0$. This result differs from the conclusion that prof. N. Perra and prof. L.E.C. Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}). They found that the minority opinion was able to survive over time, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}. However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$. In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$. In both cases the maximum length of the actual time-line is set equal to $S=20$. The difference in the results is most likely an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$. This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$. In this figure we can see that the minority opinion 0 doesn't vanish from the system over time, but instead reaches a value of $P_0(500) \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.03$ in Figure \ref{ws10000}.  % explain further, but first test WS, K=6, beta=0.01, N=1000, 20-80, PR

% Talk about effect network topologies?? See paper + simulation with N = 10.000 and N = 100.000 --> mention possible discrepancies and effect of network size??. Having a high clustering coefficient and/or a clear community structure plays no role in the e

@@

@@

@@

@@

The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity) look qualitatively the same. Similarly, the results for the Watts-Strogatz model and the stochastic Watts-Strogatz model agree with each other as well. This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is mainly the clustering that drives these differences.\\
There is no visible effect for having a clear community structure. This can be clearly seen in Figure \ref{sbm} where the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure) are compared. There is (almost) no difference in the results for a high and a low modularity.\\

@@

Networks with a high clustering coefficient (WS and SBM-WS) behave completely different. For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}). 
% remake plots?? because for PR minority opinion vanishes does why more echo chambers??? 

>>>>

@@

    \label{ws_20-80_op}

@@

    \label{sbm-ws_20-80_op}

@@

From Figure \ref{ev_op_20_80}, it is clear that the minority opinion 0 eventually ceases to exist, $P_0(500) \sim 0$. This is certainly the case for the low clustering networks (ER and SBM). The high clustering coefficient in the WS and SBM-WS networks seems to slightly hamper this vanishing of the minority opinion with an average value of $P_0(500) \sim 0.007$ for the SBM-WS network (Figure \ref{sbm-ws_20-80_op}) and an average value of $P_0(500) \sim 0.009$ for the WS network (Figure \ref{ws_20-80_op}).\\ 
This result differs from the conclusion that prof. N. Perra and prof. L.E.C. Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}). They found that the minority opinion was able to survive over time for all the network models, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}. However, the results that they obtained are for networks with a total number of nodes equal to $N = 10^5$ and an average degree around $\left<k\right> \sim 6$. In this thesis, on the other hand, results are obtained for $N=10^3$ and an average degree around $\left<k\right> \sim 10$. In both cases the maximum length of the actual time-line is set equal to $S=20$. The difference in the results is most likely due to an interplay between the network size $N$, the average degree $\left<k\right>$ and the maximum length of the actual time-line $S$. This can for example be seen in Figure \ref{ev_op_WS_k=6}, which shows the results for the evolution of the fraction of opinion 0 for an initial $20/80$ opinion distribution and using the PR method; Figure \ref{ws1000} displays the results for a WS model with $N=10^3,\ K=6$ and $\beta=0.01$; Figure \ref{ws10000} shows the results for a WS model with $N=10^4,\ K=6$ and $\beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the maximum length of the actual time-line is $S=20$. In this figure we can see that the minority opinion 0 reaches higher final values of $P_0(500)$ compared to the case depicted in Figure \ref{ws_20-80_op}, with $P_0(500 \sim 0.027$ in Figure \ref{ws1000} and a value of $P_0(500) \sim 0.03$ in Figure \ref{ws10000}.

@@

The results for the Erd\H{o}s-R\'{e}nyi model and the stochastic block model (both the high and low modularity case) look qualitatively the same. Similarly, the results for the Watts-Strogatz model and the stochastic Watts-Strogatz block model agree with each other as well. This difference in results between ER/SBM (negligible clustering) and WS/SBM-WS (high clustering) suggests that it is mainly the clustering that drives these differences.\\
There are no visible effects of having a clear community structure. This can be clearly seen in Figure \ref{sbm} where the results for a stochastic block model with a high modularity (clear community structure) and a low modularity (negligible community structure) are compared. There is (almost) no difference in the results for a high and a low modularity.\\

@@

Networks with a high clustering coefficient (WS and SBM-WS) behave completely different. For both the REC and PR method the high clustering enhances the formation of echo chambers around the minority opinion 0 (clearly visible in the spike around $\left<P_1^{\text{nn}}\right> = 0$ in both Figure \ref{ws} and Figure \ref{sbm-ws}). An important note must be made here: the absolute value of neighborhoods with all the nodes carrying the minority opinion 0 at $t=0$ is equal to zero in the case of a WS network in Figure \ref{ws} and in the case of a SBM-WS network in Figure \ref{sbm-ws}. This leads to a normalized average value of the echo chambers, which is obtained by dividing the value at $t=500$ by the corresponding value at $t=0$, equal to infinity. In Figures \ref{ws} and \ref{sbm-ws}, the spikes around zero aren't actually reaching infinity, but are cut off around values equal to 100/180.\\
This enhancing of echo chambers around the minority opinion in case of networks with a high clustering coefficient is remarkable, especially in the case of the PR filtering algorithm. Remember from Figure \ref{ev_op_20_80} that in case of the PR method the minority opinion 0 decreases to less than $1 \%$ of the entire population. Even though the total fraction of the minority opinion strongly decreases, the high clustering forces the remaining nodes with this minority opinion to group together and form opinion clusters. The normalized value of the formation of echo chambers around the majority opinion reaches values approximately equal to $F_N(\left<P_1^{\text{nn}}\right> = 1) = 1$ when using the REC filtering algorithm. This means that the average fraction of nodes with all their neighbors carrying the majority opinion 1 remains the same before and after the time evolution. In case of the PR method this value is even slightly higher than 1, indicating a small increase in the number of majority opinion echo chambers after the time evolution. 
% talk about small spike around 1 for PR method??
